---
title: "bm practice2"
author: "Ruihan Zhang"
date: "2023-03-08"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
library(mgcv)
library(earth)
```

```{r}
college_data = read_csv("D:/lenovo/Documents/College.csv", show_col_types = FALSE) %>% 
  janitor::clean_names() %>% 
  na.omit()
```

```{r}
set.seed(2023)
college_mtx <- model.matrix(outstate ~.,college_data)[,-1]
trainRows <- createDataPartition(y = college_data$outstate, p = 0.8, list = FALSE)
# design matrix
train_data <- college_data[trainRows,]
test_data <- college_data[-trainRows,]
x <- train_data %>% dplyr::select(-outstate)
y <- train_data$outstate
```

## a)

```{r}
perc.grid <- seq(from = 0, to = 70, by = 1)
fit.ss <- smooth.spline(train_data$perc_alumni, train_data$outstate)
ss.df <- fit.ss$df
ss.df
pred.ss <- predict(fit.ss,
                   x = perc.grid)
pred.ss.df <- data.frame(pred = pred.ss$y,
                         perc = perc.grid)
p <- ggplot(data = train_data, aes(x = perc_alumni, y = outstate)) +
     geom_point(color = rgb(.2, .4, .2, .5))
p + geom_line(aes(x = perc, y = pred), data = pred.ss.df,
          color = rgb(.8, .1, .1, 1)) + theme_bw()
```

The degree of freedom obtained by generalized cross-validation is `r ss.df`. The fitted spline curve performs a straight line as shown.

```{r}
perc.grid <- seq(from = 0, to = 70, by = 1)
ps <- list()
for (i in 1:63){
    fit_ <- smooth.spline(train_data$perc_alumni, train_data$outstate, df = i+1 )
    df_ <- fit_$df
    if(df_< i){
        next
    }
    pred_<- predict(fit_,
                   x = perc.grid)
    preddf_ <- data.frame(pred = pred_$y,
                         perc = perc.grid,
                         df = rep(i+1, length(pred_$y)))
    ps <- rbind(ps, preddf_)
}
p.mass <- ggplot(data = train_data, aes(x = perc_alumni, y = outstate)) +
    geom_point(color = rgb(.2, .4, .2, .5))
p.mass + geom_line(aes(x = perc, y = pred, group = df, color = df), data = ps[which(ps$df <= 15),]) + theme_bw()
```

The plot above shows the fitted smoothing spline models using perc.alumni as the only predictor of `Outstate` for a range of degree of freedom from 2 to 15. As the `df` increase, the fitted spline curve become more and more wiggly.
## b)

```{r }
gam.m1 <- gam(outstate~ s(apps) + s(accept) + s(enroll)+ s(top10perc)+s(top25perc)+s(f_undergrad)+s(p_undergrad)+s(room_board)+s(books)+s(personal)+s(ph_d)+s(terminal)+s(s_f_ratio)+s(perc_alumni)+s(expend)+s(grad_rate) , data = train_data)
summary(gam.m1)
```

The GAM model included all the predictors. The model summary shows that some `edf` equals to `1`, which means these predictors may have linear relationships with `outstate`.

```{r problem_c_plot}
plot.gam(gam.m1, pages = 4)
```
From the plot we can also see some linear trend in `apps`, `enroll`, `p_undergrad`, `personal` and `terminal`.

```{r }
gam.test.predict <- predict.gam(gam.m1, newdata = test_data, type = "response")
gam.test.mse <- (RMSE(gam.test.predict, test_data$outstate))^2
```
The test MSE of the GAM model is `r gam.test.mse`.
## c)

Train the MARS model using all predictors.
```{r,warning=FALSE}
ctrl1 = trainControl(method = "cv", number = 10)
mars_grid <- expand.grid(degree = 1:3, nprune = 2:15)
set.seed(2023)
mars.fit <- train(train_data %>% dplyr::select(-outstate), train_data$outstate, method = "earth", tuneGrid = mars_grid, trControl = ctrl1)
```